# Practical Machine Learning - Course Project Writeup

In this project, data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants was used. Participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways.Analysis was done on the provided data to predict the activity an individual performs. For this analysis, caret and randomForest package was used and answers for 20 test data cases were predicted.

```{r}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)

set.seed(2048)
options(warn=-1)
```

## Cleaning Data
Training and Test data was downloaded from the links provided in the assignment. Data contained "#DIV/0!" for some values. Such values were replaced by NA.

```{r}
training_data <- read.csv("pml-training.csv", na.strings=c("#DIV/0!") )
evaluation_data <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!") )
```


## Modeling


All columns 8 to the end were casted to be numeric. Empty columns were not taken into consideration for Feature set Selectcion. Columns like username, timestamps and windows were also excluded from Feature set selection. Feature set data was further modelled to have model_data.

```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}

feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```

Training and Test data was classified from this model data.

```{r}
idx <- createDataPartition(y=model_data$classe, p=0.70, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```

Five random forests were build with 150 trees each.

Data was partitioned by the `classe` variable to ensure the training set and test set contain examples of each class. 70% of the training data was allocated to the training set and the remainder for the validation set.
```{r}
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

Error reports were provided for both training and test data.
```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

```

Conclusions 
-----------

Confusion matrix for this model is very accurate and all the answers computed for submissed were correct.  So, it can be concluded that the random forest algorithm appears to perform very well for predicting activities from accelerometers measurements.

References: 
-----------
 http://groupware.les.inf.puc-rio.br/har. 

